---
title: 'FIN4080: HW1'
author: "Si Chen"
date: "Sep 2, 2019"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Q1: The first question 

### a. You could insert new code section by shortcut ctrl+alt+I

### b. When printing dataframe, only print first 5 lines unless more lines are required

```{r}
# your R code here
x=matrix(seq(1,50),25,2)
print(head(x))
```
### c. The following is an example of showing your code, but not running the code

```{r,echo=T,eval=F}
# WRDS Connection 
library(RPostgres)
wrds = dbConnect(Postgres(),
                  host='wrds-pgdata.wharton.upenn.edu',
                  port=9737,
                  user='your username',
                  password='your password',
                  dbname='wrds',
                  sslmode='require')
```

### d. The following example is hiding your code but running the code, check your enviornment for variable y

```{r,echo=F}
y=10
```

### e. Here's an example of showing how to write math (Optinal)

* You may hand in your math part in RMarkdown or by handwriting as attached pdf files. 

$\dfrac{P_{t}}{P_{t-1}}$

$\alpha=10$

* More information about writing math [https://en.wikibooks.org/wiki/LaTeX/Mathematics](https://en.wikibooks.org/wiki/LaTeX/Mathematics)

### f. Insert pictures
```{r}
plot(cars, pch = 18)
```

### g. More Information on using R Markdown
[https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf](https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf)

## Q2: 

## Q3: 




